---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.height = 6
)
```

# iscoCrosswalks

<!-- badges: start -->
<!-- badges: end -->

The goal of iscoCrosswalks is to map indicators and raw counts from the 
International Standard Classification of Occupations (ISCO) to the
Standard Occupational Classification (SOC) System, and vice versa.

## Installation

You can install the development version of iscoCrosswalks from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("eworx-org/crosswalkR")
```

## Example

This is a basic example which shows you how to translate CEDEFOPs
"Importance of foundation skills" indicator given in ISCO(2008) to SOC(2010)
classification:

- For further details visit [CEDEFOPs skills
intelligence tool](https://www.cedefop.europa.eu/en/tools/skills-intelligence)

```{r example}
library(iscoCrosswalks)
```

```{r, echo=FALSE}
library(ggplot2)
foundation_skills[, Skill := gsub(" skills", "", Skill)]
dat <- foundation_skills[order(Skill, -Value)][, head(.SD, 6), by = "Skill"]
ggplot(dat, aes(x = Occupations, y = Value, fill = Skill)) +
  geom_col(alpha = 0.8, width = 0.85) +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(expand = c(0, 0.1)) +
  coord_flip() +
  facet_grid(rows = vars(Skill), scales = "free_y", switch = "y", space = "free_y") +
  labs(
    title = "Which foundation skills are very important for jobs?",
    subtitle = "Highest level needed for doing their job - ISCO 2008",
    caption = "Source: Cedefop's European skills and jobs survey",
    y = "% of jobs for which foundation skills are importan"
  ) +
  theme_minimal(base_family = "Roboto Condensed") +
  theme(
    plot.margin = margin(0.5, 0.5, 0.5, 0.5, unit = "cm"),
    plot.title = element_text(size = 15, face = "bold"),
    strip.text.y = element_text(angle = 270, face = "bold"),
    strip.placement = "outside",
    axis.title.x = element_text(margin = margin(t = 0.5, b = 0.5, unit = "cm")),
    axis.title.y = element_blank(),
    axis.text = element_text(size = 10),
    legend.position = "none",
    panel.grid.major.y = element_blank(),
    )
```

The percentage of jobs where foundation skills (literacy, numeracy, ICT,
and foreign languages) are highly crucial for doing the work is shown in this
indicator. Based on the findings of Cedefop's European survey of skills and
jobs.

The Skills Foundation Indicator is exposed also in `iscoCrosswalks` as an
example data-set. It consists of three variables

- `Occupations`
- `Skill`
- `Value`

To perform the transformation, we've added a third column with the
`preferredLabel` from the ISCO taxonomy. In the R terminal, type `isco` to
access the desired labels. Manual entry of preferred labels is suggested for
small data. See also the R package
[labourR](https://cran.r-project.org/web/packages/labourR) for automating the
occupations coding, in case of big data-sets.

Inspecting the indicator,

```{r}
knitr::kable(foundation_skills[seq(1 , nrow(foundation_skills), by = 5), ])
```

To translate the indicator to SOC classification, `iscoCrosswalks` has two
mandatory column names. Namely, `job` and `value` standing for the preferred
labels of the taxonomy and the value of the indicator respectively. 

Thus, in our example we should rename `preferredLabel` to `job`, and `Value` to
`value`.

```{r}
data.table::setnames(foundation_skills,
                     c("preferredLabel", "Value"),
                     c("job", "value"))
```

Next, using the `isco_soc_crosswalk()` function we can translate the values to
the desired taxonomy. The parameter `brkd_cols` accepts a vector that indicates
other columns used for grouping. 

Also, since this is a composite score we set `indicator = TRUE` to use `mean`
value. If raw counts are given then we set `indicator = FALSE` to aggregate 
the units of the hierarchy.

```{r}
soc_foundation_skills <- isco_soc_crosswalk(foundation_skills,
                                            brkd_cols = "Skill",
                                            isco_lvl = 1,
                                            soc_lvl = "soc_1",
                                            indicator = TRUE)
```

In the following we visualize the top 6 occupations by Skill, of the projected
indicator to the SOC taxonomy.

```{r, echo=F}
library(ggplot2)
library(data.table)
soc_foundation_skills[, Occupations := gsub(" Occupations", "", soc_label)]
soc_foundation_skills[, Skill := gsub(" skills", "", Skill)]
dat <- soc_foundation_skills[order(Skill, -value)][, head(.SD, 6), by = "Skill"]
ggplot(dat, aes(x = Occupations, y = value, fill = Skill)) +
  geom_col(alpha = 0.8, width = 0.85) +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(expand = c(0, 0.1)) +
  coord_flip() +
  facet_grid(rows = vars(Skill), scales = "free_y", switch = "y", space = "free_y") +
  labs(
    title = "Which foundation skills are very important for jobs?",
    subtitle = "Highest level needed for doing their job - SOC 2010",
    caption = "Source: Cedefop's European skills and jobs survey transformed with iscoCrosswalks",
    y = "% of jobs for which foundation skills are importan"
  ) +
  theme_minimal(base_family = "Roboto Condensed") +
  theme(
    plot.margin = margin(0.5, 0.5, 0.5, 0.5, unit = "cm"),
    plot.title = element_text(size = 15, face = "bold"),
    strip.text.y = element_text(angle = 270, face = "bold"),
    strip.placement = "outside",
    axis.title.x = element_text(margin = margin(t = 0.5, b = 0.5, unit = "cm")),
    axis.title.y = element_blank(),
    axis.text = element_text(size = 10),
    legend.position = "none",
    panel.grid.major.y = element_blank(),
    )
```

If the reverse process is required, use the `soc_isco_crosswalk()` function.
