---
title: "isco-crosswalks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{isco-crosswalks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.height = 2
)
library(knitr)
```

```{r setup}
library(iscoCrosswalks)
library(data.table)
```

# Introduction

Eurostat, CEDEFOP, and other EU agencies have been looking into using the ESCO classification to
link job–worker characteristics like skill importance and level ratings to European jobs. O*NET is a
reliable source of occupational data for the European labour market because of its strong theoretical
and empirical basis and the similarities between the European and US economic systems.
Drawing on O\*NET necessitates mapping from one classification to another since the O\*NET
database classifies jobs differently than the EU. This is accomplished via a “concordance” (also known
as a “crosswalk” or “correspondence table”). We propose the a publicly accessible software for
methodologically transparent concordances, by constructing an R-package that we hope will greatly
reduce the expenses and time required to perform an approximate mapping between the two
categorization systems.

ESCO began evaluating several options to tying skills to professions in the EU. The introduction of the
Skills and Competencies Taxonomy, knowledge descriptors, which provides consistent nomenclature
and enhances communication of occupational information in the EU, was the first stage in that
process. One obvious source to consider as we evaluate the various options to utilize the ESCO
Taxonomy is the US O\*NET system. Indeed, to assess the skills and competencies of different
segments of the EU labour market, various governmental and non-governmental organizations in the
EU have built concordances between the US System of Occupational Classification (SOC) and ESCO.
The problem is that there is no clear one-to-one mapping from the EU ESCO to the US SOC utilized in
O\*NET for most jobs. As a result, many concordances have concentrated solely on occupations with
clear one-to-one correspondences. Furthermore, these concordances are frequently proprietary,
with little information or explanation about how they were created, making them inaccessible to
others and limiting any critique of the approach adopted.

Understanding the variations in how jobs are organized in the US and the EU is necessary to realize
the necessity for a more granular concordance. While the SOC and the ESCO have similar goals, there
are significant distinctions in the organization and information included in occupational profiles. The
US SOC has 867 6-digit classifications for statistics reporting (e.g., number of people employed or
average wage). The EU ESCO, on the other hand, employs the ISCO 4-digit categorization, which has
436 unit groups, and adds an extra layer of information of approximately 3000 occupational
distinctions. Both systems group occupations according to a four-level hierarchy, with ESCO adding
an extra degree of detail.

# Crosswalks

We build an algorithm that conducts approximate matching between the ISCO and SOC classifications
using concordances provided by the Institute for Structural Research and Faculty of Economics,
University of Warsaw. The crosswalks offer a complete step-by-step mapping of O\*NET data to 
ISCO88 and ISCO-08 coding using an expanded version of SOC-00 and SOC-10 coding (used for most
European data, including EU LFS). We propose a mapping method based on the aforementioned
research that converts measurements to the smallest possible unit of the target taxonomy, and then
performs an aggregation/estimate to the appropriate degree of detail. In case of raw-count data
the `sum` by breakdown group is used, else the `mean` value.

## Example (ISCO-SOC)

## Example

This is a basic example which shows you how to translate CEDEFOPs
"Importance of foundation skills" indicator given in ISCO(2008) to SOC(2010)
classification:

- For further details visit [CEDEFOPs skills
intelligence tool](https://www.cedefop.europa.eu/en/tools/skills-intelligence)

```{r example}
library(iscoCrosswalks)
```

The percentage of jobs where foundation skills (literacy, numeracy, ICT,
and foreign languages) are highly crucial for doing the work is shown in this
indicator. It is based on the findings of Cedefop's European survey of skills
and jobs.

The Skills Foundation Indicator is exposed also in `iscoCrosswalks` as an
example data-set. It consists of three variables

- `Occupations`
- `Skill`
- `Value`

To perform the transformation, we've added a third column with the
`preferredLabel` from the ISCO taxonomy. In the R terminal, type `isco` to
access the desired labels. Manual entry of preferred labels is suggested for
small data. See also the R package
[labourR](https://cran.r-project.org/web/packages/labourR) for automating the
occupations coding, in case of big data-sets.

Inspecting the indicator,

```{r}
kable(foundation_skills[seq(1 , nrow(foundation_skills), by = 5), ])
```

To translate the indicator to SOC classification, `iscoCrosswalks` has two
mandatory column names. Namely, `job` and `value` standing for the preferred
labels of the taxonomy and the value of the indicator respectively. 

Thus, we rename `preferredLabel` to `job`, and `Value` to `value`.

```{r}
data.table::setnames(foundation_skills,
                     c("preferredLabel", "Value"),
                     c("job", "value"))
```

The `isco_soc_crosswalk()` function can translate the values to the desired
taxonomy. The parameter `brkd_cols` accepts a vector that indicates
other columns used for grouping. 

Also, since this is a composite score we set `indicator = TRUE` to use `mean`
value. Instead, if raw counts are given then we set `indicator = FALSE` to
aggregate the units of the hierarchy.

```{r}
soc_foundation_skills <- isco_soc_crosswalk(foundation_skills,
                                            brkd_cols = "Skill",
                                            isco_lvl = 1,
                                            soc_lvl = "soc_1",
                                            indicator = TRUE)
```

In the following table we visualize the top 6 occupations by Skill, of the
projected indicator to the SOC taxonomy.

```{r}
soc_foundation_skills[, Occupations := gsub(" Occupations", "", soc_label)]
soc_foundation_skills[, Skill := gsub(" skills", "", Skill)]
dat <- soc_foundation_skills[order(Skill, -value)][, head(.SD, 6), by = "Skill"]
kable(dat)
```

If the reverse process is required, use the `soc_isco_crosswalk()` function. The
preffered labels of the taxonomy can be inspected in the included dataset 
`soc_groups`.
